---
title: "Exploratory Analysis"
author: "Cathal Sweeney (20320166)"
date: "2024-03-01"
output: html_document
---

```{r}
library(ggcorrplot)
library(car)
library(tidyverse)
library(VIM)
library(tidymodels)
library(vcd)

```






#investigating for missing data which may be present 
```{r}
NBAmodel <- read_csv("nba_with_next_seas_bmp.csv")

NBAmodel %>%
  filter(g>=15)

 
VIM::aggr(NBAmodel)

```

#visualizing categorical data 
```{r}
train_data %>%
  ggplot(aes(x = pos, y = bpm)) +
  geom_boxplot()
```
```{r}
train_data %>%
  ggplot(aes(x = team_id, y = bpm)) +
  geom_boxplot()
```
#Relationship between Categorical Data and our Outcome Variable 
#no evidence to suggest being part of a particular team makes a significant impact on a players bpm scores 



 




#investigating relationships of numerical variables in dataset with our outcome variable "Bpm" - 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(rlang)
predictor_variable <- "bpm"
explanatory_variables <- c( "ft_per_g",  "ast_per_g" , "pts_per_g",  "mp",   "trb_per_g",  "gs", "g",  "mp_per_g" , "fg_per_g",  "fg_pct" , "fg3_per_g",    "fg2_pct" , "ft_pct", "orb_per_g",  "stl_per_g", "blk_per_g",  "tov_per_g" ,"pf_per_g",  "ts_pct" , "per", "fta_per_fga_pct",  "trb_pct" ,  "stl_pct" ,"tov_pct",  "usg_pct",  "age" , "fg3a_per_fga_pct", "vorp","win_loss_pct",)
for (var in explanatory_variables) {
  plot_title <- paste("Scatterplot with lowess smoother for", var)
  p <- ggplot(NBA_screened1, aes(x = !!sym(var), y = !!sym(predictor_variable))) +
    geom_point() +
    geom_jitter(width = 0.2, height = 0.2, alpha = 0.4, color = "mediumpurple1") +
    geom_smooth(method = "loess", formula = y ~ x, color = "royalblue4") +
    labs(x = var, y = predictor_variable, title = plot_title) + 
    theme_minimal() +
    theme(plot.background = element_rect(fill = "whitesmoke"),
          panel.grid.major = element_line(color = "azure3"),
          panel.grid.minor = element_line(color = "azure3"),
          legend.position = "bottom") +
    scale_fill_brewer(palette = "Set3")
  
  print(p)
}
```

#investigating correlation between numeric variables specifically for fitting a linear regression model 
```{r}
nba_next_season_model[sample(1:nrow(nba_next_season_model), 1000),] %>%
  select_if(is.numeric) %>%
  cor() %>%
  ggcorrplot()
```

#Data Filtering 
#Removal of NA values present in the dataset 

```{r}
set.seed(222)

NBAcleanfullmodel1 <- 
  NBAmodel %>%
  select(player, next_seas_bpm ,pos, team_id,season,ft_per_g, ast_per_g, pts_per_g ,mp, bpm ,efg_pct, trb_per_g,gs,g,mp_per_g,fg_per_g,fga_per_g,fg_pct,fg3_per_g,fg3a_per_g,fg3_pct,fg2_per_g,fg2a_per_g,fg2_pct,fta_per_g,ft_pct,orb_per_g,orb_pct,drb_per_g,stl_per_g,blk_per_g,tov_per_g,pf_per_g,per,ts_pct,fta_per_fga_pct,drb_pct,trb_pct,ast_pct,stl_pct,blk_pct,tov_pct,usg_pct,ows,dws,ws,ws_per_48,obpm,dbpm,age,fg3a_per_fga_pct,vorp,mov,mov_adj,win_loss_pct) %>%
  drop_na()
```


#Filtering out data with bpm  above 15 and below -2. This is removing any outliers that still may remain. We also removed players who played less than 15 games during the season as there stats may be skewed. For instance a player who plays 1 or 2 games may have a high bpm which would not be sustainable for a season. We believe that 15 games which represents almost a quarter of the season is an accurate measurement of a players contribution.
```{r}
NBA_screened1 <- NBAcleanfullmodel1 %>%
  filter(g>=15) %>%
  filter(!(next_seas_bpm >= 15 | next_seas_bpm <= -2)) %>%
  filter(!(bpm >= 15 | bpm <= -2))
```

```{r}
NBA_screened1 %>%
  ggplot(aes(x=next_seas_bpm)) +
  geom_histogram() +
  ggtitle("Distribution of filtered next_seas_bpm")
```


