---
title: "Linear Regression"
author: "Cathal Sweeney (20320166) & Eoghan Tinney"
date: "2024-03-05"
output: html_document
---


```{r}
library(ggcorrplot)
library(car)
library(tidyverse)
library(VIM)
library(tidymodels)
library(vcd)

```

```{r}
data_split <- initial_split(NBA_screened1, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)

```

#Cross Validation - creating 10 folds for  cross validation
```{r}
library(tidyverse)
set.seed(345)
folds <- vfold_cv(train_data, v = 10)
folds
```
#variables for transformation based on exploratory analysis
#per,ts_pct,bpm,stl_pct,ws_per_48,fg2_pct,fg_pct,ft_pct,fg_pct,ft_pct,orb_per_g,pf_per_g,fta_per_fga_pct,fg3a_per_fga_pct

```{r}
# Define your tuning grid
degree_grid <- tibble(degree = seq(1, 5, by = 1))  # Degrees from 1 to 5
```
#finding the optimal ploynomial degree of each variable using the step_poly function 
```{r}
per_var <- recipe(next_seas_bpm ~ per, data = train_data) %>% step_poly(per, degree = tune())
```
#creating a workflow for each variable
```{r}
per_var_wf <- workflow() %>%
  add_recipe(per_var) %>%
  add_model(linear_reg())
```
#tunes each variable workflow to find the optimal degree of polynomial
```{r}
tuned_model <- per_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```
#Variable 2 true shooting percentage 
```{r}
ts_pct_var <- recipe(next_seas_bpm ~ ts_pct, data = train_data) %>% step_poly(ts_pct, degree = tune())
```

```{r}
ts_pct_var_wf <- workflow() %>%
  add_recipe(ts_pct_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- ts_pct_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```
#bpm variable 3 

```{r}
bpm_var <- recipe(next_seas_bpm ~ bpm, data = train_data) %>% step_poly(bpm, degree = tune())
```

```{r}
bpm_var_wf <- workflow() %>%
  add_recipe(bpm_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- bpm_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

#variable 4 stl_pct
```{r}
stl_pct_var <- recipe(next_seas_bpm ~ stl_pct, data = train_data) %>% step_poly(stl_pct, degree = tune())
```

```{r}
stl_pct_var_wf <- workflow() %>%
  add_recipe(stl_pct_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- stl_pct_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

#variable 5 ws_per_48

```{r}
ws_per_48_var <- recipe(next_seas_bpm ~ ws_per_48, data = train_data) %>% step_poly(ws_per_48, degree = tune())
```

```{r}
ws_per_48_var_wf <- workflow() %>%
  add_recipe(ws_per_48_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- ws_per_48_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

#variable 6 fg2_pct
```{r}
fg2_pct_var <- recipe(next_seas_bpm ~ fg2_pct, data = train_data) %>% step_poly(fg2_pct, degree = tune())
```

```{r}
fg2_pct_var_wf <- workflow() %>%
  add_recipe(fg2_pct_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- fg2_pct_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

#variable 7 fg_pct
```{r}
fg_pct_var <- recipe(next_seas_bpm ~ fg_pct, data = train_data) %>% step_poly(fg_pct, degree = tune())
```

```{r}
fg_pct_var_wf <- workflow() %>%
  add_recipe(fg_pct_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- fg_pct_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

#variable 8 ft_pct
```{r}
ft_pct_var <- recipe(next_seas_bpm ~ ft_pct, data = train_data) %>% step_poly(ft_pct, degree = tune())
```

```{r}
ft_pct_var_wf <- workflow() %>%
  add_recipe(ft_pct_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- ft_pct_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

#variable 9 orb_per_g
```{r}
orb_per_g_var <- recipe(next_seas_bpm ~ orb_per_g, data = train_data) %>% step_poly(orb_per_g, degree = tune())
```

```{r}
orb_per_g_var_wf <- workflow() %>%
  add_recipe(orb_per_g_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- orb_per_g_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```
##variable 10 pf_per_g
```{r}
pf_per_g_var <- recipe(next_seas_bpm ~ pf_per_g, data = train_data) %>% step_poly(pf_per_g, degree = tune())
```

```{r}
pf_per_g_var_wf <- workflow() %>%
  add_recipe(pf_per_g_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- pf_per_g_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

###variable 11 fta_per_fga_pct
```{r}
fta_per_fga_pct_var <- recipe(next_seas_bpm ~ fta_per_fga_pct, data = train_data) %>% step_poly(fta_per_fga_pct, degree = tune())
```

```{r}
fta_per_fga_pct_var_wf <- workflow() %>%
  add_recipe(fta_per_fga_pct_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- fta_per_fga_pct_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

##fg3a_per_fga_pct
```{r}
fg3a_per_fga_pct_var <- recipe(next_seas_bpm ~ fg3a_per_fga_pct, data = train_data) %>% step_poly(fg3a_per_fga_pct, degree = tune())
```

```{r}
fg3a_per_fga_pct_var_wf <- workflow() %>%
  add_recipe(fg3a_per_fga_pct_var) %>%
  add_model(linear_reg())
```

```{r}
tuned_model <- fg3a_per_fga_pct_var_wf %>%
  tune_grid(resamples = folds, grid = degree_grid, control = control_grid(verbose = TRUE))
```

```{r}
collect_metrics(tuned_model)
```


```{r}
show_best(tuned_model)
```

#post transformation



```{r}
lr_mod <- 
  linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")
```

```{r}
# Combine all recipes into one
full_recipe <- 
  recipe(next_seas_bpm ~ usg_pct + age + fg3a_per_fga_pct + fg2_pct + trb_pct + ft_pct+ fta_per_fga_pct +  ft_per_g + ast_per_g + pts_per_g + mp  + trb_per_g + gs + g + mp_per_g  + fg_pct + fg3_per_g   + fg2_per_g   + orb_per_g + orb_pct + stl_per_g + blk_per_g + tov_per_g + pf_per_g + per + ts_pct + fta_per_fga_pct + stl_pct + tov_pct + bpm +pos , data = train_data)
```


#Creating a Workflow for a linear regression Model
```{r}
lr_wf_new <- 
  workflow() %>%
  add_model(lr_mod) %>%
  add_recipe(full_recipe )
```


```{r}
set.seed(456)
lr_wf_new_fit <- 
  lr_wf_new %>%
  fit(data = train_data)
lr_wf_new_fit
```
#performing k-fold cross-validation
```{r}
linear_reg_new_fit <- lr_wf_new_fit %>%
  fit_resamples(folds)
```
#gathering the mean and standard error cross validation rmse 

```{r}
collect_metrics(linear_reg_new_fit)
```









